---
title: "ANOVA"
author: "Michał Kędra"
output: html_document
---

```{r, warning=FALSE, echo =FALSE, message =FALSE  }
library(ggplot2)
library(psych)     
library(lawstat)   
library(dunn.test)
library(tseries)
library(readxl)

```

### Introduction
ANOVA is a technique used to analise the differences among group means in a sample. For example, with ANOVA we can group people by age and see if mean of iq within all groups is equal. The null hypothesis for ANOVA is that all population means are exactly equal.

we will check if birth weight is affected by age of the mother and race. Birth weight is the body weight of a baby at its birth. The average birth weight in babies of European heritage is 3.5 kilograms. But to begin, let's see, how data looks.


### Seeing data


```{r}
births<-read_excel("births.xls")
tail(births)

```
As we can see we have many variables in data set, but we will focus on belove described ones:

* __BWT__ - birth weight in grams
* __AGE__ -  mother age, when baby was borned
* __RACE__ - skin color (1 = white, 2 = black, 3 = other)



```{r}
summary(births$AGE)
```

Now we will change __AGE__  to  a binary variable, 1 if mothers age is under or equal 23 and 0 if age is below 23.
```{r}
births$AGE[births$AGE < 23]<- 0
births$AGE[births$AGE >= 23 ]<- 1
```

Let's see types of our variables
```{r}
str(births)
```

AGE and RACE are change to a categorical variable:
```{r}
births$RACE<- factor(births$RACE, labels = c("white", "black", "other"))
births$AGE<- factor(births$AGE, labels = c("0-23", "23+"))
```

Boxplots are showing the distribution of __BWT__ (birth weight)  among groups.

1. Grouped by race.
```{r}
ggplot(births)+ geom_boxplot(aes(births$RACE, births$BWT, color= births$RACE))
```

We can see slight diffrences beetwen groups

2. Grouped by age

```{r, echo = FALSE}
ggplot(births)+ geom_boxplot(aes(births$AGE, births$BWT, color= births$AGE))


```

As we can see  boxplots are almost same, expect min and max value

3. Grouped by race and age.

```{r, echo = FALSE}
ggplot(births)+ geom_boxplot(aes(births$RACE, births$BWT, color= births$AGE))

```

Age makes the biggest diffrnence among black race.

Statistisc in groups:
```{r}
describeBy(births$BWT, births$RACE, mat = TRUE)
```

```{r, echo = FALSE}
describeBy(births$BWT, births$AGE, mat = TRUE)
```

To do ANOVA we need normality distribution among all groups.
It will be test with Shaprio test:

```{r}
shapiro_pvalue<-function (x){
  return (shapiro.test(x)$p.value)
}
tapply(births$BWT, births[,c(1,3)], shapiro_pvalue)

```
So all groups have normal distribution, and we can do anova test.

### Anova

```{r}
births_anova <- aov(BWT~AGE*RACE, data = births)
summary(births_anova)
```

```{r}
summary_anova<- summary(births_anova)
eta.age<- summary_anova[[1]][1,2]/(summary_anova[[1]][1,2]+summary_anova[[1]][4,2])
eta.race<- summary_anova[[1]][2,2]/(summary_anova[[1]][2,2]+summary_anova[[1]][4,2])
eta.race_age<- summary_anova[[1]][3,2]/(summary_anova[[1]][3,2]+summary_anova[[1]][4,2])   

```


```{r}
omega.age<- (summary_anova[[1]][1,2]-summary_anova[[1]][1,1]*summary_anova[[1]][4,3])/(summary_anova[[1]][4,3]+sum(summary_anova[[1]][,2]))

omega.race<- (summary_anova[[1]][2,2]-summary_anova[[1]][2,1]*summary_anova[[1]][4,3])/(summary_anova[[1]][4,3]+sum(summary_anova[[1]][,2]))

omega.race_age<- (summary_anova[[1]][3,2]-summary_anova[[1]][3,1]*summary_anova[[1]][4,3])/(summary_anova[[1]][4,3]+sum(summary_anova[[1]][,2]))
```

